\documentclass[11pt]{report}

% codes
\newcommand{\pele}{{\sf Pele}}
\newcommand{\pelec}{{\sf PeleC}}
\newcommand{\pelelm}{{\sf PeleLM}}
\newcommand{\pelephysics}{{\sf PelePhysics}}
\newcommand{\fuego}{{\sf Fuego}}
\newcommand{\iamr}{{\sf IAMR}}
\newcommand{\amrex}{{\sf AMReX}}
\newcommand{\craypat}{{\sf CrayPat}}

% amrex vocab
\newcommand{\cpp}{{\sf C++}}
\newcommand{\fortran}{{\sf Fortran}}
\newcommand{\omp}{{\sf OpenMP}}
\newcommand{\bx}{{\sf Box}}
\newcommand{\bxarray}{{\sf BoxArray}}
\newcommand{\fab}{{\sf FAB}}
\newcommand{\fabarray}{{\sf FabArray}}
\newcommand{\multifab}{{\sf MultiFab}}
\newcommand{\dmap}{{\sf DistributionMap}}
\newcommand{\mfiter}{{\sf MFIter}}
\newcommand{\profparser}{{\sf AMReXProfParser}}

% landscape orientation for tables
\usepackage{pdflscape} 

%------------------------------------------------------------------------------
\begin{document}

\chapter{Overview}

Completed and in-progess changes to \iamr.
\begin{enumerate}
\item
  Update initialization procedure to build EB-aware geometries
\item
  Update nodal projection
\item
  Implement EB-friendly advection scheme in \iamr. Special care needed for cut cells with small volume fraction in order to preserve conservation.
\item
  Update computation of external forcing terms
\item
  EB-ize MAC Projection
\item
  Scalar diffusion rearrangement to speed-up problems with large multi-species networks. Also required modifications in the syncronization
\item
  EB updates for diffusion.
  \begin{itemize}
  \item
    Transition away from volume-weighted diffusivities/viscosities in order to maximize use of existing \amrex\ functionality. 
  \item
    Update scalar and velocity diffusion to use EB-aware solvers
  \item
    Debugging and validation (can we use this term?) of tensor solver
  \item
    Transition tensor solve from using divU computed from temperature, mass fractions and enthalpy (i.e.\ Diffusion::compute\_divmusi) to using divU computed from the velocity field  
  \end{itemize}
\item
  Multilevel updates
  \begin{itemize}
  \item
    Enforce restriction that EB and coarse-fine boundary do not cross. This restriction is also currently embedded in \amrex\ diffusion solvers and perhaps additional solvers (projections?).
  \item
    Ensure data stuctures are created with approriate knowledge of the EB.
    Requires a detailed examination of the code because not supplying EB information does not necessarily lead to crashes, only incorrect answers.
    NavierStokesBase, NavierStokes, and Diffusion.cpp are good, but haven't touched MacProj or Projection
  \end{itemize}  
\end{enumerate}

To Do items
\begin{enumerate}
\item
  Multilevel
  \begin{itemize}
  \item
    GradP - \iamr\ goes to lengths to enure non-EB gradp is consistent across levels. Would like to check on this for gradp coming from MLMG
  \item
    create\_umac\_grown
  \item
    SyncAdvect
  \item
    MAC sync
  \item
    Diffusion Ssync, Vsync 
  \item
    Sync projection
  \end{itemize}
\end{enumerate}


\end{document}
